Endogenous Politics and \\ the Design of Trade Institutions

### 2017-01-16 draft.tex
rewrite end of sec:objfcn

The setup that I have assumed has very similar dynamics. When the cost of lobbying effort is included in government welfare as in Equation~\ref{eq:wel}, the monotonicity in $\ga$ inherent in \ref{eq:baldwin} typically breaks down. For most specifications of the $\ga(e)$ function, there is an interior optimum. If a government's objective function is closer to Expression~\ref{eq:wel} than Expression~\ref{eq:baldwin} with $e$ endogenized, the government will likely be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences. By setting a tariff cap below the lobby's optimal tariff, the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters.\footnote{There are many candidates for the government objective function that are not monotonic in political pressure. For example, in \Textcite{ethier2012}, returns to lobbying decrease as tariffs increase. In the current model, one could take government welfare to be quadratic in $\ga$ or make the weights on the terms in the objective function sum to one.}

@@

There are several additional advantages to the setup assumed here. First, it allows me to take lobbying incentives into account in the same framework typically used in the literature to examine institutional design questions so that comparisons to the existing literature are direct. Second, it is very tractable, making it easily adopted to ask how endogenous politics changes the answers to a wide range of policy questions. Third, if one prefers a non-unitary interpretation of the government, this framework in which the lobby exerts effort that determines the identity of the pivotal decision-maker within the government is more appealing than bargaining. Finally, in an environment where exogenous shocks are present, contingent lobbying effort / payments are hard to reconcile with many types of uncertainty.

If one takes $a=1$, the only difference between the \Textcite{lt} government objective function and Expression~\ref{eq:wel} is that in place of their $\Psi(e)$, I have $[\ga(e) - 1]\pi_x(\tau) - e$.\footnote{Note that this is still consistent with the \Textcite{dgh97} form, although of course I have the lobby maximizing net profits instead of making a take-it-or-leave-it offer.}	This means there's a positive cross derivative between lobbying effort and tariffs in my formulation, so that an increase in tariffs increases government welfare more the higher is lobbying effort.

>>>>

The setup that I have assumed is similar to \Textcite{lt} and produces very similar dynamics.\footnote{If one takes $a=1$, the only difference between the \Textcite{lt} government objective function and Expression~\ref{eq:wel} is that Expresion~\ref{eq:wel} replaces $\Psi(e)$ with $[\ga(e) - 1]\pi_x(\tau) - e$. This means there's a positive cross derivative between lobbying effort and tariffs in my formulation, so that an increase in tariffs increases government welfare more the higher is lobbying effort. Note that this is still consistent with the \Textcite{dgh97} form, although I have the lobby maximizing net profits instead of making a take-it-or-leave-it offer. } When the cost of lobbying effort is included in government welfare as in Equation~\ref{eq:wel}, the monotonicity in $\ga$ inherent in \ref{eq:baldwin} typically breaks down. For most specifications of the $\ga(e)$ function, there is an interior optimum. If a government's objective function is closer to Expression~\ref{eq:wel} than Expression~\ref{eq:baldwin} with $e$ endogenized, the government will likely be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences. By setting a tariff cap below the lobby's optimal tariff, the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters.\footnote{There are many candidates for the government objective function that are not monotonic in political pressure. For example, in \Textcite{ethier2012}, returns to lobbying decrease as tariffs increase. In the current model, one could take government welfare to be quadratic in $\ga$ or make the weights on the terms in the objective function sum to one.}

@@

There are several additional advantages to the setup assumed here. First, it is very tractable, making it easily adopted to ask how endogenous politics might change the answers to a wide range of policy questions. Second, it allows me to take lobbying incentives into account in the same framework typically used in the literature to examine institutional design questions so that comparisons to the existing literature are direct. Third, if one prefers a non-unitary interpretation of the government, this framework in which the lobby exerts effort that determines the identity of the pivotal decision-maker within the government is more appealing than bargaining. Finally, in an environment where exogenous shocks are present, contingent lobbying effort / payments are hard to reconcile with many types of uncertainty.

### 2017-01-16 to-do.tex
rewrite end of sec:objfcn

				\item The fact that the applied tariffs are equal to the negotiated binding is reminiscent of Maggi and Rodriguez-Clare 2007. Exactly the same dynamic is at play here: specifying trade agreement tariffs as caps instead of strong bindings keeps the lobby active during periods where the trade agreement is honored. In Appendix B.3, I analyze the model with strong bindings and show that the results would be altered in magnitude but not in spirit by assuming that the trade agreement tariffs are strong bindings instead of tariff caps. The only change to the model is that under strong bindings there would be zero lobbying effort during a trade agreement phase as the lobby would not need to exert effort to bid protection levels up to the trade agreement tariff. There would still be no disputes in equilibrium, but the lobbying constraint is easier to satisfy with strong bindings because the gap between trade war and trade agreement profits shrinks when the lobby stops exerting effort to receive the trade agreement tariff. Because lower trade agreement tariffs can be sustained under strong bindings, executive and legislators with small political economy weights prefer a strong-binding agreement and legislators with high political economy weights prefer a weak-binding agreement. A tariff cap makes the lobby's self-enforcement constraint harder to satisfy and thus requires a higher trade agreement tariff for self-enforcement. From the ex-ante point of view, strong bindings are therefore preferable. Tariff caps could be viewed as a way to commit to setting higher trade agreement tariffs and therefore as a mechanism for ensuring that rents are distributed to protectionists ex-post.
			\end{itemize}
		\item MRC07
			\begin{itemize}
				\item Need case of fixed capital ($z=0$)
				\item NB (p. 1391): optimal trade agreement if lobbies have all the bargaining power and no ex-ante lobbying is free trade
					\begin{itemize}
						\item Page 26-28 of working paper has government with all the bargaining power (and no capital mobility, but remember that the capital owners choose where to invest in the first period). Here, could be free trade but could be interior to get some rents.
					\end{itemize}		
			\end{itemize}
		\item Make argument that want this version like Baldwin-style literature for comparison
			\begin{itemize}
				\item Want to make argument that things go wrong in standard model when you insert $e$ into $\ga$ (also, when $e$ goes into $\ga$ need to subtract $e$ too, and this often results in interior optimum)
				\item This is an easier way to get the result from MRC07 / LT, can make comparisons, develop design results more easily
			\end{itemize}
	\end{itemize}

>>>>

				\item The fact that the applied tariffs are equal to the negotiated binding is reminiscent of Maggi and Rodriguez-Clare 2007. Exactly the same dynamic is at play here: specifying trade agreement tariffs as caps instead of strong bindings keeps the lobby active during periods where the trade agreement is honored. In Appendix B.3, I analyze the model with strong bindings and show that the results would be altered in magnitude but not in spirit by assuming that the trade agreement tariffs are strong bindings instead of tariff caps. The only change to the model is that under strong bindings there would be zero lobbying effort during a trade agreement phase as the lobby would not need to exert effort to bid protection levels up to the trade agreement tariff. There would still be no disputes in equilibrium, but the lobbying constraint is easier to satisfy with strong bindings because the gap between trade war and trade agreement profits shrinks when the lobby stops exerting effort to receive the trade agreement tariff. Because lower trade agreement tariffs can be sustained under strong bindings, executive and legislators with small political economy weights prefer a strong-binding agreement and legislators with high political economy weights prefer a weak-binding agreement. A tariff cap makes the lobby's self-enforcement constraint harder to satisfy and thus requires a higher trade agreement tariff for self-enforcement. From the ex-ante point of view, strong bindings are therefore preferable. Tariff caps could be viewed as a way to commit to setting higher trade agreement tariffs and therefore as a mechanism for ensuring that rents are distributed to protectionists ex-post.
			\end{itemize}
		\item MRC07
			\begin{itemize}
				\item Need case of fixed capital ($z=0$)
				\item NB (p. 1391): optimal trade agreement if lobbies have all the bargaining power and no ex-ante lobbying is free trade
					\begin{itemize}
						\item Page 26-28 of working paper has government with all the bargaining power (and no capital mobility, but remember that the capital owners choose where to invest in the first period). Here, could be free trade but could be interior to get some rents.
					\end{itemize}		
			\end{itemize}
	\end{itemize}
